Discussion Draft

### 2020-04-05 doc/Discussion.rmd
distracted by papers

Our results of greater between cell variation in females, these results are supported by human and mouse (Gruhn, Lynn Kong). [@lynn2002].

These data are also scant, between cell variance isn't often measured especially for both sexes. (given they require single cell observations for both sexes).

While the modifier models parameterize sexual dimorphism in the strength of selection which can shape the variance in a trait these types of models make predictions  for variance between individuals but not within individuals where the genetics are constant. (selection on penetrance).

In the spindle based selection model, the source of between cell variance is attributed to difference strength of checkpoints.
As mentioned above, there is sexual dimorphism in the stringency of the spindle assembly checkpoint (SAC). And these are linked to conserved features of gametogenesis.

The opposite dynamics of the purifying selection for the 'strict'  SAC in spermatocytes is relaxed selection from a 'permissive' SAC. As mentioned above a fundamental sex difference in the metaphase spindle is the presence of centrioles in spermatocytes which anchors and nucleates the microtubulue spindle at the two cell poles. While oocytes lack centrioles which results in a difuse network of microtubules with mulitple microtubule organizing centers (aMTOCs) (Schuh Ellenburg 2007). This or larger cell volumne may lead/be a potnetial source of the leaky SAC in oocytes. Leaky SAC, relaxed selection on the anaphase to metaphase transition could increase the between cell variance for crossover number

>>>>

Our results of greater between cell variation in females, these results are supported by human and mouse [@lynn2002, @gruhn2013].

While the **modifier models** parameterize sexual dimorphism in the strength of selection which can shape the variance in a trait these types of models make predictions  for variance between individuals but not within individuals where the genetics are constant.

In the **spindle based** selection model, the less efficient SAC in oocytes could lead to relaxed selection on the chromosome structure. If the SAC/checkpoint which normally acts in a stabilizing selection on the chromosome structure has a reduction in efficiency (relaxed selection), will result in increased variance in (these associated traits), not just across strains (populations), but across cells within individuals (quant-gen ref?). Oocytes at metaphase I, have an acentriolar spindle comprised of  diffuse network of microtubules with multiple microtubule organizing centers (aMTOCs) (Schuh Ellenburg 2007) which result in

<stabilization -- bi-oriented, robust SAC response, SAC acts to correctly time metaphase to anaphase transition -- only after correct alignment and biorientation of bivalents in the metaphase cell> -- spindle dynamics much different due to asymetrical diviation Mogessie et al 2018, Mogessie Schuh --actin and MT are apart of oocyte spindle.

- which might have a consequence of reducing the (reaction/signal of the SAC (anaphase promoting complexes (cite) )).  (this is another way the spindle based selection model is linked to conserved sex differences in gametogenesis (cite))

<Bennabi et al 2016 --spindle assembly inoocytes, sun and kim 2011 -- SAC and regulators, Kolano et al 2012 --error prone oocytes, Manil-Segalen -- forces through MT breaking chromsomes>
Mogessie et al 2018 -- meiotic spindle (big)review>
(what generates MT nucleation is unclear (aMTOC or chromsoomes))
